---
title: "Spanish syllable organization"
output: html_notebook
---
Run for emptying the environment
```{r}
rm(list = ls()) 
```

### Load needed packages

```{r}
library(psych)
library(xtable) #for texy tables
library(ggplot2)
library(ez)

```

### Loading data, cleaning, adding stress- and CCV-column

```{r}
vp08 <- read.table("results/data/vp08_Antonia_new_copy.csv", header = TRUE, sep=";")
vp09 <- read.table("results/data/vp09_copy.csv", header = TRUE, sep=";")

vp08$SPEAKER <- as.factor("vp08")
vp09$SPEAKER <- as.factor("vp09")
```

### Creating new dataframe for ease of computation

```{r}
easyfun <- function(df, anc = "MAXC..ms.") {
  
  df$SEGMENT <- sub("C2.+", "C2", df$SEGMENT) # substitute C2z (etc.) with only C2
  df$SEGMENT <- sub("CODA.+", "CODA", df$SEGMENT) # substitute CODAxy (etc.) with only CODA
  
  if(df$SPEAKER[1] == "vp09"){
    df <- subset(df, SOURCE != "doblen_10"&SOURCE != "polen_07") # those instances are not fully labeled
  }
  
  df <- df[c("SOURCE", "TRAJ", "SEGMENT", "GONS..ms.", "NONS..ms.", "MAXC..ms.", "NOFFS..ms.", "GOFFS..ms.", "SPEAKER")]
  df$STRESS <- 0
  df$STRESS[grepl(".*l.r.*|saplen.*", df$SOURCE )] <- 1 #all stressed items end in r, except Saplén

  df$CCV <- 0
  df$CCV[grepl(".*pl|.*bl", df$SOURCE )] <- 1
  
  nei <- aggregate(df$STRESS, list(df$SOURCE), mean)
  colnames(nei) <- c("SOURCE", "STRESS")
  #add CCV-column back since it got lost while aggregating - there surely is a better way to do this
  nei$CCV <- 0
  nei$CCV[grepl(".*pl|.*bl", nei$SOURCE )] <- 1
  
  # add RE
  nei$RE <- df[df$SEGMENT == "C2", "NOFFS..ms."]
  
  # add anchor
  nei$ANC <- df[df$SEGMENT == "CODA", anc] # could be also Nons OR spatial extremum of vowel (see paper)
  
  # add LE -> conditional on CV or CCV
  nei[nei$CCV == 1, "LE"] <- df[df$SEGMENT == "C1", "NONS..ms."]
  nei[nei$CCV == 0, "LE"] <- df[df$SEGMENT == "C2"&df$CCV == 0, "NONS..ms."]

  # add C-Center
  nei$CC <- nei$LE + ((nei$RE-nei$LE)/2)
  
  # add IPI for CCVs
  # make CCV subframe
  
  ccv_df <- df[df$CCV == 1,]
  nei$normIPI <- NA
  #C2 target(NONS) - C1 release (NOFFS) / RE - LE 
  nei[nei$CCV == 1, "normIPI"] <- (ccv_df[ccv_df$SEGMENT == "C2", "NONS..ms."] - ccv_df[ccv_df$SEGMENT == "C1", "NOFFS..ms."]) / (nei[nei$CCV == 1, "RE"] - nei[nei$CCV == 1, "LE"])
  nei$IPI <- NA
  nei[nei$CCV == 1, "IPI"] <- (ccv_df[ccv_df$SEGMENT == "C2", "NONS..ms."] - ccv_df[ccv_df$SEGMENT == "C1", "NOFFS..ms."])
  
  # add speaker column
  nei$SPEAKER <- df[1,"SPEAKER"]
  
  return(nei)
}
```
Create a new datafram that contains the relevant values for us
```{r}
short_08 <- easyfun(vp08)
short_09 <- easyfun(vp09)
# remove hablar_06 and hablar_03 -> IPI & normIPI outlier, to large & negative
short_09 <- short_09[short_09$SOURCE != "hablar_06" & short_09$SOURCE != "hablar_03", ]
new_df <- rbind(short_08, short_09)
new_df$STRESS <- as.factor(new_df$STRESS)
head(new_df)

```

### Some sanity checks:


```{r}
sum(new_df$LE < new_df$RE) # left edge is left to the right edge
sum(new_df$RE < new_df$ANC) # anchor is right to the right edge
sum(new_df$CC < new_df$ANC) # C-center is left to anchor
```
Since there are 257 rows, there is one instance where the right edge is before the left edge this instance is coplan_01 of the first speaker. I will investigate that later, but first let's just remove the instance for the time being.

```{r}
new_df <- new_df[new_df$LE < new_df$RE,]
instances <- nrow(new_df)
write(instances, file = "text/numbers/instances.txt")
instances
```
```{r}
#Add column with item names
new_df$WORD <- sub(pattern = "([^\\d]+)_\\d{2}.*",replacement = "\\1", x = new_df$SOURCE)
```

### Now calculating some relative standard deviations
```{r}
stress <- new_df[new_df$STRESS == 1, ]
nostress <- new_df[new_df$STRESS == 0, ]


relsds_fun <- function(df){
  le <- sd(df$ANC - df$LE)/mean(df$ANC - df$LE)*100
  cc <- sd(df$ANC - df$CC)/mean(df$ANC - df$CC)*100
  re <- sd(df$ANC - df$RE)/mean(df$ANC - df$RE)*100
  return(c(le, cc, re))
} 
# stressed 
  # anchor - le
s_le <- (stress$ANC - stress$LE) /mean(stress$ANC - stress$LE)*100
  # anchor - cc
s_cc <- (stress$ANC - stress$CC)/mean(stress$ANC - stress$CC)*100
  # anchor - re
s_re <- (stress$ANC - stress$RE)/mean(stress$ANC - stress$RE)*100
# unstressed
  # anchor - le
n_le <- (nostress$ANC - nostress$LE)/mean(nostress$ANC - nostress$LE)*100
  # anchor - cc
n_cc <- (nostress$ANC - nostress$CC)/mean(nostress$ANC - nostress$CC)*100
  # anchor - re
n_re <- (nostress$ANC - nostress$RE)/mean(nostress$ANC - nostress$RE)*100


rel_sds <- data.frame("Stressed_syllable" = c("yes", "no"),
           LE = c(sd(s_le), sd(n_le)),
           CC = c(sd(s_cc), sd(n_cc)),
           RE = c(sd(s_re), sd(n_re))
           ) 
rel_sds

```
```{r}
print(xtable(rel_sds, caption = "Relative standard deviations  of interval length between left edge (LE), C-Center (CC) and right edge (RE) and the anchor, respectively.\\label{tab:rel_sds}", type = "latex"), file = "text/tables/rel_sds.tex")
```
Distribtution of the normalized interval lengths (first row not stressed, second row stressed)
```{r}
par(mfrow = c(2,3))
boxplot(n_le, ylim = c(0,2))
boxplot(n_cc, ylim = c(0,2))
boxplot(n_re, ylim = c(0,2))
boxplot(s_le, ylim = c(0,2))
boxplot(s_cc, ylim = c(0,2))
boxplot(s_re, ylim = c(0,2))

```



Let's look closer

```{r}

blA <- new_df[new_df$WORD == "molar"|new_df$WORD == "hablar",]
blE <- new_df[new_df$WORD == "soler"|new_df$WORD == "tobler",]
plA <- new_df[new_df$WORD == "molar"|new_df$WORD == "coplar",]
plE <- new_df[new_df$WORD == "soler"|new_df$WORD == "saplen",]
bla <- new_df[new_df$WORD == "molan"|new_df$WORD == "hablan",]
ble <- new_df[new_df$WORD == "polen"|new_df$WORD == "doblen",]
pla <- new_df[new_df$WORD == "molan"|new_df$WORD == "coplan",]
ple <- new_df[new_df$WORD == "polen"|new_df$WORD == "soplen",]

x <- list(blA, blE, plA, plE, bla, ble, pla, ple)
all_pairs <- data.frame(Landmark = c("LE", "CC", "RE"),
                        sapply(x, relsds_fun))
names(all_pairs) <- c("Landmark", "blA","blE", "plA", "plE", "bla", "ble", "pla", "ple")

print(xtable(all_pairs, caption = "Relative standard deviations of interval length between left edge (LE), C-Center (CC) and right edge (RE) and the anchor, respectively. Broken down by syllable type; capitalization of letters indicates stress.\\label{tab:all_pairs}", type = "latex"), file = "text/tables/rel_sds_pairs.tex")
all_pairs
```


### Let's look at the IPI

```{r}
boxplot(new_df$IPI, outline = FALSE)

#new_df$CCV <- as.factor(new_df$CCV)

plot <- ggplot(data = subset(new_df, !is.na(IPI)), aes(x = STRESS, y = normIPI)) +
  geom_violin() + 
  stat_summary(fun = "mean",
               geom = "point",
               aes(color = "Mean")) +
  stat_summary(fun = "median",
               geom = "point",
               aes(color = "Median")) +
  #scale_x_discrete(labels=WORD)+
  theme(axis.text.x = element_text(angle = 20))+
  scale_colour_manual(values = c("red", "blue"), # Colors
                      name = "")

plot + ggtitle("IPI") +
  xlab("Word") + ylab("Normalized IPI")

ggsave("text/figures/IPI_per_word_vio.png",
       height = 4,
       width = 6.4,
       units = "in",
       dpi = 600)

t.test(new_df[new_df$STRESS == 0, "IPI"], new_df[new_df$STRESS == 1, "IPI"])
plot 


```

How does stress lengthen the syllables?
```{r}
new_df$LEN <- new_df$ANC - new_df$LE

plot2 <- ggplot(data = new_df, aes(x = STRESS, y = LEN)) +
  geom_violin() + 
  stat_summary(fun = "mean",
               geom = "point",
               aes(color = "Mean")) +
  stat_summary(fun = "median",
               geom = "point",
               aes(color = "Median")) +
  #scale_x_discrete(labels=WORD)+
  #theme(axis.text.x = element_text(angle = 20))+
  scale_colour_manual(values = c("red", "blue"), # Colors
                      name = "")

plot2 +  xlab("Stress") + ylab("Syllable length in ms") #ggtitle("Violing plot of syllable length depending on stress") +

ggsave("text/figures/syll_length.png",
       height = 4,
       width = 6.4,
       units = "in",
       dpi = 600)
```

```{r}
len_unstr <- new_df[new_df$STRESS == 0,"LEN"]
len_str <- new_df[new_df$STRESS == 1,"LEN"]

ttest <- t.test(len_str, len_unstr, alternative = "greater")
write(ttest$p.value, file = "text/numbers/ttest_length_pvalue.txt")
cwrite(ttest$method, file = "text/numbers/ttest_length_method.txt")
write(round(ttest$estimate[1]), file = "text/numbers/ttest_length_estx.txt")
write(round(ttest$estimate[2]), file = "text/numbers/ttest_length_esty.txt")

```



